Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge

### 2025-06-11 paper/main.tex
fixed propositions

The following is immediately obtained as a corollary.

\begin{proposition}
For a conditional independence oracle
 $\mathcal{D}_G$ and an acyclic $G$-compatible ancestral oracle $\mathcal{A}_H$, 
$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)
=H^+ \supseteq G^+$.
\label{prop:weakexpand}
\end{proposition}

When using experts that generate cyclic graphs, edges are pruned to 
break these cycles, which can lead to a more sparse result.

\begin{proposition}
For a conditional independence oracle
 $\mathcal{D}_G$ and a $G$-compatible ancestral oracle $\mathcal{A}_H$ where
$H$ contains at least one cycle,
let $\tilde{G}=\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)$
Then $H^+ \supsetneq \tilde{G} \supseteq G$. 
\label{prop:weakexpand:cyclic}
\end{proposition}

\begin{proof}
All edges in the skeleton of $G$ are added during $\textsc{Expand}$ with the 
correct orientation and cannot be removed by $\textsc{FixCycles}$. Therefore, the result
is a supergraph of $G$. If cycles are created during $\textsc{Expand}$, then every 
cycle contains at least one edge that is not the skeleton of $G$, otherwise $G$ itself
would be cyclic. At least one edge of every cycle is therefore removed by
 $\textsc{FixCycles}$, making the result acyclic again.
\end{proof}

>>>>

When using experts that do not always correctly detect the absence of 
causal relationships, the resulting graph can get larger but also smaller,
if this leads to the occurrence of cycles that need to be broken.

\begin{proposition}
For a conditional independence oracle
 $\mathcal{D}_G$ and a $G$-compatible ancestral oracle $\mathcal{A}_H$, 
let $\tilde{G}=\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)$
Then $H^+ \subseteq \tilde{G} \supseteq G$. If $H$ contains at least one cycle, then
$H^+ \subsetneq \tilde{G}$.
\label{prop:weakexpand:cyclic}
\end{proposition}

\begin{proof}
All edges in the skeleton of $G$ are added during $\textsc{Expand}$ with the 
correct orientation and cannot be removed by $\textsc{FixCycles}$. Therefore, the result
is a supergraph of $G$. If cycles are created during $\textsc{Expand}$, then every 
cycle contains at least one edge that is not in the skeleton of $G$, otherwise $G$ itself
would be cyclic. At least one edge of every cycle is therefore removed by
 $\textsc{FixCycles}$, making the result acyclic again. Since the removed edges need
to be in $H^+$, the first inclusion is strict for cyclic $H$.
\end{proof}
